---
title: "Activity: Street or Road?"
author: "Mysha Nihal"
output: html_notebook
---

## Load in the data

entire `Addresses` data set. Explain how your code works line by line.

```{r}

#load libraries

library(magrittr)
library(readr)
library(dplyr)
library(ggplot2)
#load in the data here

Addresses <- read_csv("https://mdbeckman.github.io/dcSupplement/data/street-addresses.csv")

```

## Recreate Solved Example using `sample(50)`

```{r}

Sample <-
  Addresses %>%
    sample_n(size = 50)
head(Sample)

pattern <- 'PO'

Matches <-
  Sample %>%
    filter(grepl(pattern = pattern, address))
head(Matches)

Dont <-
  Sample %>%
    filter(!grepl(pattern = 'PO', address))
head(Dont)

pattern <- 'BOX\\s+(\\d+)'

Matches <-
  Sample %>%
    filter( grepl(pattern, address))
head(Matches)

Dont <-
  Sample %>%
    filter(!grepl(pattern, address))
head(Dont)

BoxNumbers <-
  Sample %>%
    filter(grepl(pattern, address)) %>%
    tidyr::extract(address, into = 'boxnum', regex = pattern)
head(BoxNumbers)

```

## Recreate Solved Example using the entire `Address` data set

```{r}

pattern <- 'PO'

Matches <-
  Addresses %>%
    filter(grepl(pattern = pattern, address))
head(Matches)

Dont <-
  Addresses %>%
    filter(!grepl(pattern = 'PO', address))
head(Dont)

pattern <- 'BOX\\s+(\\d+)'

Matches <-
  Addresses %>%
    filter( grepl(pattern, address))
head(Matches)

Dont <-
  Addresses %>%
    filter(!grepl(pattern, address))
head(Dont)

BoxNumbers <-
  Addresses %>%
    filter(grepl(pattern, address)) %>%
    tidyr::extract(address, into = 'boxnum', regex = pattern)
head(BoxNumbers)

```

#### Additional Question: How many addresses are a PO BOX (including NCSU BOX)?

```{r}

Matches <- 
  Addresses %>%
  filter(grepl(pattern = "PO BOX|NCSU BOX", address))

count(Matches)

```

## Back to the Streets: Your Turn

#### Part 1: Explain (in english) each line of the following code chunk

```{r}

# pattern <- "(ST|RD|ROAD)" = pattern that matches the strings "ST", "RD", or "ROAD" is assigned to the variable 'pattern'.

# LeftOvers <- = filtered rows assigned to the 'LeftOvers' data frame.

#  Addresses %>% = takes the Addresses data frame as the input.

#  filter( !grepl(pattern, address), = excludes rows where the 'address' column contains "ST", "RD", or "ROAD" using the 'grepl()'.

#          !grepl("\\sAPT|UNIT\\s[\\d]+$", address), = excludes rows where the 'address' column contains "APT" or "UNIT" followed by a space and one or more digits using the 'grepl()' function.

#          !grepl(" BOX ", address) = excludes rows where the address column contains the exact phrase " BOX " using the grepl() function.


```

#### Part 2: Implement your method on the entire Addresses data set. Explain how your code works line by line.

```{r}

pattern2 <- "(AVE|BLVD|PL)"
NewAddresses <- Addresses %>%
filter(grepl(pattern2, address))

show(NewAddresses)

# pattern2 <- "(AVE|BLVD|PL)" = assigns the pattern "(AVE|BLVD|PL)" to the variable 'pattern2'.

# NewAddresses <- Addresses %>% = creates a new data frame called 'NewAddresses' by filtering the Addresses data frame.

#filter(grepl(pattern2, address)) = filter the address that contains the pattern.

#show(NewAddresses) = show the new data frame.

```

#### Additional Question 1: Present your result from part 2 by providing a table in descending order of popularity for the street name endings you found

```{r}

EndingCounts <- table(NewAddresses$address) %>%
  as.data.frame() %>%
  arrange(desc(Freq))

show(EndingCounts)

```

#### Additional Question 2: Present your result by using ggplot to a bar chart in descending order of popularity for the street name endings you found previously.

Note: make sure your order street names according to popularity, not alphabetically.

```{r}

ggplot(data = EndingCounts, aes(x = reorder(Var1, -Freq), y = Freq)) +
  geom_bar(fill = "red", stat = "identity") +
  labs(x = "Street Name Endings", y = "Frequency") +
  ggtitle("Popularity of Street Name Endings") +
  theme_bw()
  
```

#### Additional Question 3: Comment on any patterns you see

'RD', and 'ST' are the most common street types.
